---
title: "EDA - FE - OLS"
author: "Sijia"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: sandstone
    toc_float: true
    code_folding: hide
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{css format, echo = FALSE}
.center-leaflet {
  display: flex;
  justify-content: center;
}

```

```{r setup, include=FALSE, quiet = TRUE, results = 'hide', echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE, fig.show = "asis",
    fig.align = "center")
knitr::opts_knit$set(root.dir = ("~/Documents/Upenn/Practicum/2025/results"))

library(sf)
library(ggplot2)
library(geojsonio)
library(dplyr)
library(tidyverse)
library(tidycensus)
library(googleway)
library(osmdata)
library(ggforce)
library(geosphere)
library(ggtext)
library(showtext) 
library(sysfonts) 
library(grid)
library(lwgeom)
library(modelsummary)
library(car)
library(broom) 
library(stargazer)
library(units)
library(broom)
library(kableExtra)
library(corrplot)
library(patchwork)
library(rsample)
library(Metrics)
library(knitr)
library(tibble)
library(leaflet)
library(RColorBrewer)
```

# Data Wrangling
## Data Set up

```{r read_data, message=FALSE, warning=FALSE, quiet=TRUE, results='hide'}
property <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/property/opa_properties_public2.geojson") %>%
  st_transform('EPSG:2272')

property2 <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/property/property2.geojson") %>%
  st_transform('EPSG:2272')

property_park <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/property/property_park.geojson")%>%
  st_transform('EPSG:2272')

property2$distance_to_nearest_park <- property_park$distance_to_nearest_park

studyarea1 <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/studyarea/StudyArea.shp")%>%
  st_transform('EPSG:2272')

I676 <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/studyarea/I_676.shp")

studyarea_tract <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/studyarea/StudyArea_blockgroup_tract.shp")
philly_blockgroup <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/Philly_blockgroup/Philly_blockgroup.shp")

nhood <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/philly_neighborhoods.geojson")

landuse <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/landuse_clip/Land_Use_ClipLayer.shp")

I676_2 <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/studyarea/discontinuity.shp")%>%
  st_transform('EPSG:2272')
cpi_data <- st_read("/Users/jiajiabao/Documents/Upenn/Practicum/2025/Data/property/cpi_data.csv")

commercial<- st_read("~/Documents/Upenn/Practicum/2025/Data/Commercial_Corridors.geojson")%>%
  st_transform('EPSG:2272')
studyarea_north <- st_read("~/Documents/Upenn/Practicum/2025/Data/studyarea/studyarea_north.shp") %>%
  st_transform('EPSG:2272')

studyarea_south <- st_read("~/Documents/Upenn/Practicum/2025/Data/studyarea/studyarea_south.shp") %>%
  st_transform('EPSG:2272')

property_final <- st_read("~/Documents/Upenn/Practicum/2025/results/property_final.geojson")%>%
  st_transform('EPSG:2272')

```

## Data Processing
### property data
```{r message=FALSE, warning=FALSE, quiet = TRUE, results='hide'}
#### clean data for city wise - sale price
colnames(cpi_data) <- c("half1", "half2", "cpi_date")
cpi_long <- cpi_data %>%
  mutate(
    Jan = `half1`, Feb = `half1`, Mar = `half1`, Apr = `half1`, May = `half1`, Jun = `half1`,
    Jul = `half2`, Aug = `half2`, Sep = `half2`, Oct = `half2`, Nov = `half2`, Dec = `half2`
  ) %>%
  select(-half1, -half2) %>%
  pivot_longer(cols = -cpi_date, names_to = "Month", values_to = "CPI") %>%
  mutate(
    Month = match(Month, month.abb),
    date = as.Date(paste(cpi_date, Month, "01", sep = "-"))  
  ) %>%
  select(date, CPI)

property_philly <-property%>%
  mutate(sale_date = as.Date(sale_date, format = "%Y-%m-%d")) %>%
  filter(sale_date >= as.Date("2020-01-01") & sale_date <= as.Date("2024-12-31"))%>%
  filter(!is.na(sale_price), !is.na(total_livable_area))%>%
  mutate(
    sale_month = floor_date(sale_date, "month"))%>%
  left_join(cpi_long, by = c("sale_month" = "date"))%>%
  mutate(CPI = as.numeric(CPI))%>%
  mutate(adj_sale_price = sale_price * (340.331/CPI))%>%
  select(-sale_month, -recording_date,  -other_building, 
         -assessment_date, -mailing_address_2, mailing_city_state, mailing_care_of)

#mean_price <- mean(property_philly$adj_sale_price, na.rm = TRUE)
#sd_price <- sd(property_philly$adj_sale_price, na.rm = TRUE)
#extreme_outlier <- mean_price + 5 * sd_price
price_filtered_philly <- property_philly %>%
  filter(adj_sale_price >= 100)

q_philly <- quantile(price_filtered_philly$adj_sale_price, probs = c(0.05, 0.95), na.rm = TRUE)

property_philly2 <- property_philly %>%
  filter(adj_sale_price >= q_philly[1], adj_sale_price <= q_philly[2])

#property_philly <- property_philly%>%
#  filter(adj_sale_price > 30000, adj_sale_price < extreme_outlier) %>%
#  filter(total_livable_area >= 100)

property_philly <- st_transform(property_philly2, crs = 2272)

#st_write(property_philly, "property_philly.geojson", driver = "GeoJSON", delete_dsn = TRUE)

#### clean data for study area - sale price 
#study area intersection
property_studyarea <- st_intersection(property, studyarea1)

#sale price clean up and inflation adjusted
colnames(cpi_data) <- c("half1", "half2", "cpi_date")
cpi_long <- cpi_data %>%
  mutate(
    Jan = `half1`, Feb = `half1`, Mar = `half1`, Apr = `half1`, May = `half1`, Jun = `half1`,
    Jul = `half2`, Aug = `half2`, Sep = `half2`, Oct = `half2`, Nov = `half2`, Dec = `half2`
  ) %>%
  select(-half1, -half2) %>%
  pivot_longer(cols = -cpi_date, names_to = "Month", values_to = "CPI") %>%
  mutate(
    Month = match(Month, month.abb),
    date = as.Date(paste(cpi_date, Month, "01", sep = "-"))  
  ) %>%
  select(date, CPI)

property_data <-property_studyarea%>%
  mutate(sale_date = as.Date(sale_date, format = "%Y-%m-%d")) %>%
  filter(!is.na(sale_price))%>%
  filter(sale_date <= as.Date("2024-12-31"))%>%
  mutate(
    sale_date = as.Date(sale_date, format = "%Y-%m-%d"), 
    sale_month = floor_date(sale_date, "month"))%>%
  left_join(cpi_long, by = c("sale_month" = "date"))%>%
  mutate(CPI = as.numeric(CPI))%>%
  mutate(adj_sale_price = sale_price * (340.331/CPI)) %>%
  select(-sale_month,-recording_date,  -other_building, 
         -assessment_date, -mailing_address_2, mailing_city_state, mailing_care_of, -state_code)%>%
  relocate(adj_sale_price, sale_price, .after = 2)

#mean_price <- mean(property_data$adj_sale_price, na.rm = TRUE)
#sd_price <- sd(property_data$adj_sale_price, na.rm = TRUE)
#extreme_outlier <- mean_price + 5 * sd_price

#property_data <- property_data %>%
#  filter(adj_sale_price > 30000, adj_sale_price < extreme_outlier)

#q <- quantile(property_data$adj_sale_price, probs = c(0.05, 0.98), na.rm = TRUE)
#property_data2 <- property_data %>%
#  filter(adj_sale_price >= 30000, adj_sale_price <= q[2])
price_filtered <- property_data %>%
  filter(adj_sale_price >= 100)

q <- quantile(price_filtered$adj_sale_price, probs = c(0.05, 0.95), na.rm = TRUE)

property_data2 <- property_data %>%
  filter(adj_sale_price >= q[1], adj_sale_price <= q[2])

property1 <- st_transform(property_data2, crs = 2272)

#join attributes - I676(NS) - nhood1 (neighborhoods in study area)
property1 <- rbind(
    property1 %>% st_intersection(studyarea_north["geometry"]) %>%
      mutate(I676_NS = "north"),
    property1 %>% st_intersection(studyarea_south["geometry"]) %>%
      mutate(I676_NS = "south")
  )
nhood1 <- st_intersection(nhood, studyarea1)

#Features clean up for future FE and modeling
## Zoning
property1 <- property1%>%
  select(1:2, zoning, location, everything())

property1 <- property1 %>%
  mutate(zoning = case_when(
    location == "1310-16 VINE ST" ~ "CMX4",   
    location == "252-56 N 13TH ST" ~ "CMX4",
    location == "255-57 N BROAD ST" ~ "CMX4",
    location == "451 N 12TH ST" ~ "CMX3",
    location == "453 N 12TH ST" ~"CMX3",
    location == "447 N 12TH ST" ~ "CMX3",
    location == "1108 BUTTONWOOD ST" ~ "CMX3",
    location == "1104 BUTTONWOOD ST" ~ "CMX3",
    location == "256-60 N MARVINE ST" ~ "CMX4",
    location == "820 VINE ST" ~ "CMX4",
    TRUE ~ zoning                        
  ))

## central air
target_locations_N <- c(
                      "1232 SUMMER ST",	
                      "1030 BRANDYWINE ST",
                      "214 N 12TH ST",
                      "1021 SPRING ST")

property1 <- property1 %>%
  mutate(
    central_air = ifelse(location %in% target_locations_N & central_air == "N", "Y", central_air)
  )
## year built - clean up values of 1 and delete the other rows which are parking plot properties that have no year_built data
valid_locations <- c(
  "305-09 N 12TH ST",
  "1008-16 WOOD ST",
  "1004-06 WOOD ST",
  "211 N CAMAC ST",
  "231-33 N BROAD ST",
  "225-27 N 12TH ST"
)
property1 <- property1 %>%
  filter(location != "711-39 SPRING GARDEN ST") %>%
  mutate(year_built = as.numeric(year_built)) %>%
  mutate(
    year_built = case_when(
      location == "305-09 N 12TH ST" & year_built < 1750 ~ 1950,
      location == "1008-16 WOOD ST" & year_built < 1750 ~ 1978,
      location == "1004-06 WOOD ST" & year_built < 1750 ~ 1920,
      location == "211 N CAMAC ST" & year_built < 1750 ~ 1900,
      location == "231-33 N BROAD ST" & year_built < 1750 ~ 2020,
      location == "225-27 N 12TH ST" & year_built < 1750 ~ 2024,
      
      !is.na(year_built) & year_built < 1750 &
        !(location %in% c(
          "305-09 N 12TH ST",
          "1008-16 WOOD ST",
          "1004-06 WOOD ST",
          "211 N CAMAC ST",
          "231-33 N BROAD ST",
          "225-27 N 12TH ST"
        )) ~ NA_real_,
      TRUE ~ year_built
    )
  )
st_write(property_philly, "property_philly.geojson", driver = "GeoJSON", delete_dsn = TRUE)
```


```{r message=FALSE, warning=FALSE}
property_data_wgs <- st_transform(property_data, crs = 4326)

filtered_out <- property_data_wgs %>%
  filter(adj_sale_price < q[1] | adj_sale_price > q[2]) %>%
  mutate(group = "Filtered Out Property")

filtered_kept <- property_data_wgs %>%
  filter(adj_sale_price >= q[1] & adj_sale_price <= q[2]) %>%
  mutate(group = "Included Property")

filtered_out_with_zoning <- filtered_out %>%
  filter(!is.na(zoning))

filtered_out_na_zoning <- filtered_out %>%
  filter(is.na(zoning)) %>%
  mutate(group = "Filtered Out (Unknown Zoning)")

zoning_levels <- unique(filtered_out_with_zoning$zoning)
n_colors <- length(zoning_levels)

color_vec <- colorRampPalette(brewer.pal(8, "Dark2"))(n_colors)
zoning_pal <- colorFactor(palette = color_vec, domain = zoning_levels)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%

  addCircleMarkers(
    data = filtered_out_with_zoning,
    radius = 3,
    stroke = FALSE,
    fillOpacity = 0.8,
    color = ~zoning_pal(zoning),
    label = ~paste0("Price: $", round(adj_sale_price, 0), "<br>Zoning: ", zoning),
    group = "Filtered Out Property",
    labelOptions = labelOptions(direction = "auto")
  ) %>%

  addCircleMarkers(
    data = filtered_out_na_zoning,
    radius = 2,
    stroke = FALSE,
    fillOpacity = 0.6,
    color = "#d3d3d3",
    label = ~paste0("Price: $", round(adj_sale_price, 0), "<br>Zoning: Unknown"),
    group = "Filtered Out (Unknown Zoning)",
    labelOptions = labelOptions(direction = "auto")
  ) %>%

  addCircleMarkers(
  data = filtered_kept,
  radius = 2,
  stroke = FALSE,
  fillOpacity = 1,      
  color = "black",       
  label = ~paste0("Price: $", round(adj_sale_price, 0)),
  group = "Included Property",
  labelOptions = labelOptions(direction = "auto")
  ) %>%

  addLegend(
    "bottomleft",
    pal = zoning_pal,
    values = filtered_out_with_zoning$zoning,
    title = "Zoning Code (Filtered Properties)"
  ) %>%

  addLayersControl(
    overlayGroups = c("Filtered Out Property", "Filtered Out (Unknown Zoning)", "Included Property"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


### Census Data
```{r message=FALSE, warning=FALSE, quiet = TRUE, results='hide'}
### The method here for pushing data is diffenrent from the one we usually use cause the there is some issues when downloading the geometry data

# census block group
census_api_key("", install = TRUE, overwrite = TRUE)
acs_variable_list.2023 <- load_variables(2023, "acs5", cache = TRUE)

local_shapefile <- "~/Documents/Upenn/Practicum/2025/Data/census/tl_2023_42_bg"
bg_pa <- st_read(local_shapefile, quiet = TRUE)
bg_philly <- bg_pa %>%
  filter(COUNTYFP == "101") %>%
  st_transform(2272) %>%
  select(GEOID, geometry)

block2023_attr <- get_acs(
  geography = "block group",
  variables = c("B19013_001E", "B25058_001E"),
  year = 2023,
  state = 42,
  county = 101,
  geometry = FALSE
)

block2023_attr <- block2023_attr %>%
  select(GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  rename(
    MedHHInc = B19013_001,
    MedRent = B25058_001
  )

block2023 <- left_join(bg_philly, block2023_attr, by = "GEOID")

studyarea_tract_df <- as.data.frame(studyarea_tract) %>% select(-geometry)
block2023_studyarea <- block2023 %>% inner_join(studyarea_tract_df, by = "GEOID")

#census tract
tract_pa <- st_read("~/Documents/Upenn/Practicum/2025/Data/census/tl_2023_42_tract", quiet = TRUE)

tract_philly <- tract_pa %>%
  filter(COUNTYFP == "101") %>%
  st_transform(2272) %>%
  select(GEOID, geometry)

tract_attr <- get_acs(
  geography = "tract",
  variables = c("B25026_001E", "B15001_050E", "B15001_009E","B06012_002E"),
  year = 2023, state = 42, county = 101, geometry = FALSE
)

tract_attr_wide <- tract_attr %>%
  select(GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  rename(
    TotalPop = B25026_001,
    FemaleBachelors = B15001_050,
    MaleBachelors = B15001_009,
    TotalPoverty = B06012_002
  )

tract2023 <- left_join(tract_philly, tract_attr_wide, by = "GEOID")
studyarea_tract_df <- as.data.frame(studyarea_tract) %>% select(-geometry)

tract2023 <- tract2023 %>%
  mutate(GEOID_trimmed = as.numeric(str_sub(GEOID, 7, 9)))

studyarea_tract_df <- studyarea_tract_df %>%
  mutate(GEOID_trimmed = as.numeric(str_sub(TRACT, 2, 4)))

tract2023_studyarea <- tract2023 %>%
  inner_join(studyarea_tract_df, by = "GEOID_trimmed")
tract_studyarea <- st_intersection(studyarea1, tract2023_studyarea)

# Property1_property2_census_join
property1 <- property1 %>%
  mutate(census_tract = as.numeric(census_tract))
tract_studyarea_selected <- tract_studyarea %>%
  select(GEOID_trimmed, TotalPoverty, MaleBachelors, FemaleBachelors, TotalPop) %>%
  st_drop_geometry()  

property1_blockgroup <- st_join(property1, block2023_studyarea %>% select(GEOID, MedHHInc, MedRent), 
                            join = st_intersects)

property1_geometry <- st_geometry(property1)

tract_studyarea_selected <- tract_studyarea_selected %>%
  distinct(GEOID_trimmed, .keep_all = TRUE)
property1 <- property1_blockgroup %>%
  st_drop_geometry() %>% 
  left_join(tract_studyarea_selected, by = c("census_tract" = "GEOID_trimmed"))

property1 <- property1 %>%
  mutate(objectid = as.character(objectid))

property2 <- property2 %>%
  mutate(objectid = as.character(objectid))

property_combined <- property1 %>%
  left_join(property2, by = "objectid") %>%
  rename(sale_price = sale_price.y)%>%
  select(-sale_price.x)
```


### Third Party Data 
```{r Sijia POI, message=FALSE, warning=FALSE, quiet=TRUE, results='hide', eval = FALSE}
# POI-restaurant-drink-dessert
api_key <- ""
keywords <- c("restaurant", "dim sum restaurant", "kitchen", "desert","tea", "bubble tea","matcha", "herbal tea")
place_types <- c("restaurant", "cafe", "bakery")

studyarea_proj <- st_transform(block2023_studyarea, 2272)

grid_pts <- st_make_grid(studyarea_proj, cellsize = 2000, what = "centers")
grid_pts <- st_sf(geometry = grid_pts) %>%
  st_filter(st_buffer(studyarea_proj, 1000)) 

grid_pts_wgs <- st_transform(grid_pts, 4326)
coords <- st_coordinates(grid_pts_wgs)

all_responses <- list()

for (pt_idx in seq_len(nrow(coords))) {
  location <- coords[pt_idx, c(2,1)]  # lat, lng
  cat(" Querying center point", pt_idx, "at", paste(location, collapse = ", "), "\n")
  
  for (ptype in place_types) {
    for (kw in keywords) {
      Sys.sleep(5)  
      cat("🔎", kw, "in", ptype, "\n")

      response <- google_places(
        keyword = kw,
        location = location,
        radius = 2000,  
        place_type = ptype,
        key = api_key
      )

      if (is.null(response[["results"]]) || length(response[["results"]]) == 0) {
        cat("No results for:", kw, "in", ptype, "\n")
        next
      }

      page_token <- response$next_page_token
      results_list <- list(response[["results"]])
      i <- 1

      while (!is.null(page_token)) {
        Sys.sleep(10)
        response_next <- google_places(
          keyword = kw,
          location = location,
          radius = 2000,
          place_type = ptype,
          key = api_key,
          page_token = page_token
        )

        if (is.null(response_next[["results"]]) || length(response_next[["results"]]) == 0) break

        results_list[[i + 1]] <- response_next[["results"]]
        page_token <- response_next$next_page_token
        i <- i + 1
      }

      all_responses[[paste(kw, ptype, pt_idx, sep = "_")]] <- bind_rows(results_list) %>%
        filter(!is.na(place_id)) %>%
        distinct(place_id, .keep_all = TRUE)
    }
  }
}

all_responses <- all_responses[lengths(all_responses) > 0]
restaurant_results <- bind_rows(all_responses) %>%
  mutate(lat = geometry$location$lat, lng = geometry$location$lng) %>%
  select(lat, lng, name, place_id, rating) %>%
  distinct(place_id, .keep_all = TRUE)

restaurant_sf <- st_as_sf(restaurant_results, coords = c("lng", "lat"), crs = 4326)
restaurant_proj <- st_transform(restaurant_sf, 2272)

studyarea_buffer <- st_buffer(studyarea_proj, 2000)
res_poi <- st_filter(restaurant_proj, studyarea_buffer)
radius_circles <- st_buffer(st_transform(grid_pts, 2272), dist = 2000)

# POI - grocery
keywords <- c("grocery store", "supermarket", "asian grocery", "chinese supermarket",
              "korean market", "japanese grocery", "international market", "food market")

place_types <- c("supermarket", "grocery_or_supermarket", "convenience_store", "store")

studyarea_proj <- st_transform(block2023_studyarea, 2272)

grid_pts <- st_make_grid(studyarea_proj, cellsize = 2000, what = "centers")
grid_pts <- st_sf(geometry = grid_pts) %>%
  st_filter(st_buffer(studyarea_proj, 1000))

grid_pts_wgs <- st_transform(grid_pts, 4326)
coords <- st_coordinates(grid_pts_wgs)

all_responses <- list()

for (pt_idx in seq_len(nrow(coords))) {
  location <- coords[pt_idx, c(2,1)]  # lat, lng
  cat(" Querying center", pt_idx, "at", paste(location, collapse = ", "), "\n")
  
  for (ptype in place_types) {
    for (kw in keywords) {
      Sys.sleep(5)  
      cat("🔎", kw, "in", ptype, "\n")

      response <- google_places(
        keyword = kw,
        location = location,
        radius = 2000,
        place_type = ptype,
        key = api_key
      )

      if (is.null(response[["results"]]) || length(response[["results"]]) == 0) next

      page_token <- response$next_page_token
      results_list <- list(response[["results"]])
      i <- 1

      while (!is.null(page_token)) {
        Sys.sleep(10)
        response_next <- google_places(
          keyword = kw,
          location = location,
          radius = 2000,
          place_type = ptype,
          key = api_key,
          page_token = page_token
        )
        if (is.null(response_next[["results"]]) || length(response_next[["results"]]) == 0) break

        results_list[[i + 1]] <- response_next[["results"]]
        page_token <- response_next$next_page_token
        i <- i + 1
      }

      all_responses[[paste(kw, ptype, pt_idx, sep = "_")]] <- bind_rows(results_list) %>%
        filter(!is.na(place_id)) %>%
        distinct(place_id, .keep_all = TRUE)
    }
  }
}

all_responses <- all_responses[lengths(all_responses) > 0]

grocery_results <- bind_rows(all_responses) %>%
  mutate(lat = geometry$location$lat,
         lng = geometry$location$lng) %>%
  select(lat, lng, name, place_id, rating) %>%
  distinct(place_id, .keep_all = TRUE)

grocery_sf <- st_as_sf(grocery_results, coords = c("lng", "lat"), crs = 4326)
grocery_proj <- st_transform(grocery_sf, 2272)

studyarea_buffer <- st_buffer(studyarea_proj, 2000)
grocery <- st_filter(grocery_proj, studyarea_buffer)

# property join poi
property_sf <- st_as_sf(property_combined, sf_column_name = "geometry")
property_proj <- st_transform(property_sf, 2272)
grocery_proj <- st_transform(grocery, 2272)
restaurant_proj <- st_transform(res_poi, 2272)

grocery_idx <- st_nearest_feature(property_proj, grocery_proj)

property_proj$nearest_grocery_name <- grocery_proj$name[grocery_idx]
property_proj$nearest_grocery_dist <- st_distance(
  property_proj,
  grocery_proj[grocery_idx, ],
  by_element = TRUE
)
restaurant_idx <- st_nearest_feature(property_proj, restaurant_proj)
property_proj$nearest_restaurant_name <- restaurant_proj$name[restaurant_idx]
property_proj$nearest_restaurant_dist <- st_distance(
  property_proj,
  restaurant_proj[restaurant_idx, ],
  by_element = TRUE
)
property_proj$nearest_grocery_dist_m <- as.numeric(property_proj$nearest_grocery_dist)
property_proj$nearest_restaurant_dist_m <- as.numeric(property_proj$nearest_restaurant_dist)
clean_property_poi_distances <- function(property_sf) {
  property_sf %>%
    mutate(
      nearest_grocery_dist_m = as.numeric(nearest_grocery_dist),
      nearest_restaurant_dist_m = as.numeric(nearest_restaurant_dist)) %>%
    select(
      -nearest_grocery_dist,
      -nearest_restaurant_dist
    )
}
property_final <- clean_property_poi_distances(property_proj)
st_write(property_final, "property_final.geojson", driver = "GeoJSON", delete_dsn = TRUE)
```


# Exploratory Data Analysis
## Highway Impacts
### Linear Relationship: Sale Price vs. Distance to I-676 by Zoning
```{r highway_effects_zoning, message=FALSE, warning=FALSE}
property_filtered <- property_final
property_filtered <- st_as_sf(property_filtered, sf_column_name = "geometry")

avg_price_zoning <- property_filtered %>%
  st_drop_geometry() %>%  
  group_by(zoning) %>%
  summarise(avg_price_zoning = mean(adj_sale_price, na.rm = TRUE)) %>%
  ungroup()

property_filtered <- property_filtered %>%
  left_join(avg_price_zoning, by = "zoning")%>%
  filter(location != "711-39 SPRING GARDEN ST")

property_filtered <- property_filtered %>%
  mutate(zoning_group = case_when(
    zoning %in% c("CMX5", "ICMX") ~ "High-Density Commercial",
    zoning %in% c("IRMX") ~ "Residential Industrial",
    zoning %in% c("RM1", "RM4", "RMX3", "RSA5") ~ "Residential",
    zoning %in% c("CMX2.5", "CMX2", "CMX1") ~ "Neighborhood Commercial",
    zoning %in% c("CMX3", "CMX4") ~ "Community Commercial",
    TRUE ~ "Other"
  ))

avg_price_group <- property_filtered %>%
  st_drop_geometry() %>%
  group_by(zoning_group) %>%
  summarise(avg_adj_sale_price = mean(adj_sale_price, na.rm = TRUE)) %>%
  ungroup()

property_filtered <- property_filtered %>%
  left_join(avg_price_group, by = "zoning_group")


ggplot(property_filtered %>% filter(!is.na(zoning_group) & zoning_group != "Other"), 
       aes(x = log1p(distance_to_I676), y = log1p(adj_sale_price))) +
  geom_point(alpha = 0.5, size = 0.5, color = "#1B8370") + 
  geom_smooth(method = "lm", se = FALSE, color = "#E4572E") +
  facet_wrap(~ zoning_group, scales = "free") + 
  theme_minimal() +
  labs(title = "                      Zoning wise Relationship Between Distance to I-676 and Sales Price", 
       x = "Distance to I-676", 
       y = "Sale Price")
```

### Linear Relationship: Sale Price vs. Distance to I-676 by Neighborhood

```{r highway_effects_nhood, message=FALSE, warning=FALSE}
property_with_nhood <- st_join(property_filtered, nhood1, join = st_intersects)

nhood_summary <- property_with_nhood %>%
  group_by(NAME) %>%
  summarise(
    avg_sale_price = mean(adj_sale_price, na.rm = TRUE),
    avg_distance = mean(distance_to_I676, na.rm = TRUE)
  ) %>%
  ungroup()
nhood_map <- st_join(nhood1, nhood_summary, by = "NAME")

property_logansquare_split <- property_filtered %>%
  mutate(
    nhoods_name = case_when(
      GEOID %in% c("421010003003") ~ "Logan Square", 
      GEOID %in% c("421010125013", "421010125014") ~ NA_character_, 
      str_to_title(str_replace_all(nhoods_name, "_", " ")) == "Center City" ~ "Center City East (Near City Hall)", 
      TRUE ~ str_to_title(str_replace_all(nhoods_name, "_", " "))
    )
  ) %>%
  filter(!is.na(nhoods_name))

avg_price_nhoods <- property_logansquare_split %>%
  st_drop_geometry() %>%
  group_by(nhoods_name) %>%
  summarise(avg_sale_price = mean(adj_sale_price, na.rm = TRUE)) %>%
  ungroup()

highlight_nhoods <- c( "Callowhill", "West Poplar")
green_nhoods <- c("Center City East (Near City Hall)", "Chinatown", "Old City", "Logan Square")

property_logansquare_split$nhoods_name <- factor(
  property_logansquare_split$nhoods_name,
  levels = c(highlight_nhoods, green_nhoods)
)
ggplot(property_logansquare_split %>% filter(!is.na(nhoods_name)), 
       aes(x = log1p(distance_to_I676), y = log1p(adj_sale_price), 
           color = ifelse(nhoods_name %in% highlight_nhoods, "North Area", "South Area"))) +
  geom_point(alpha = 0.5, size = 0.5) +  
  geom_smooth(method = "lm", se = FALSE, color = "black") +  
  facet_wrap(~ nhoods_name, scales = "free", nrow = 2) +  
  scale_color_manual(values = c("North Area" = "#E4572E", "South Area" = "#1B8370")) +  
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.justification = c(0.5, -0.2)
  ) +
  labs(
        title = "                   Neighborhood wise Relationship Between Distance to I-676 and Sales Price", 
       x = "Distance to I-676", 
       y = "Sale Price",
    color = "Neighborhood"
  )
```

### Correlation Matrix
```{r correlation matrix, message=FALSE, warning=FALSE}
lm1_property <- property_final %>%
  st_drop_geometry() %>%  
  select(
    adj_sale_price,
    distance_to_I676,
    distance_to_city_hall,
    distance_to_nearest_transit,
    distance_to_nearest_bikelane,
    distance_to_nearest_hospital,
    distance_to_nearest_school,
    nearest_restaurant_dist_m,
    nearest_grocery_dist_m,
    distance_to_nearest_park,
    distance_to_nearest_water,
    total_livable_area,
    interior_condition,
    exterior_condition,
    number_of_bathrooms,
    number_of_bedrooms,
    garage_spaces,
    general_construction,
    crime_nn5
  ) %>%
  mutate(across(everything(), as.numeric)) %>%
  na.omit()

cor_matrix <- cor(lm1_property, use = "complete.obs")

var_labels <- c(
  "adj_sale_price" = "Sale Price",
  "distance_to_I676" = "Distance to Highway I-676",
  "distance_to_city_hall" = "Distance to City Hall",
  "distance_to_nearest_transit" = "Distance to Transit",
  "distance_to_nearest_bikelane" = "Distance to Bikelane",
  "distance_to_nearest_school" = "Distance to School",
  "nearest_restaurant_dist_m" = "Distance to Restaurant",
  "nearest_grocery_dist_m" = "Distance to Grocery",
  "distance_to_nearest_hospital" = "Distance to Hospital",
  "distance_to_nearest_park" = "Distance to Park",
  "distance_to_nearest_water" = "Distance to Water",
  "total_livable_area" = "Total Livable Area",
  "interior_condition" = "Interior Condition",
  "exterior_condition" = "Exterior Condition",
  "number_of_bathrooms" = "# of Bathrooms",
  "number_of_bedrooms" = "# of Bedrooms",
  "garage_spaces" = "# of Garage Spaces",
  "crime_nn5" = "Crime Rate",
  "general_construction" = "General Construction"
)

colnames(cor_matrix) <- var_labels[colnames(cor_matrix)]
rownames(cor_matrix) <- var_labels[rownames(cor_matrix)]

corrplot(
  cor_matrix,
  method = "color",
  col = colorRampPalette(c("#1B8370", "white", "#E46726"))(200),
  type = "lower",
  tl.pos = "lt",
  tl.col = "black",
  tl.cex = 0.6,
  mar = c(0, 0, 5, 0)
)

title(
  main = "Correlation Across Key Numeric Features",
  cex.main = 1.5,
  font.main = 2,
  line = 0.3,   
  adj = 0.6  
)
```


## Feature Engineering
```{r FE countinued , quiet=TRUE, results='hide'}
median_bedroom <- median(as.numeric(as.character(property_final$number_of_bedrooms)), na.rm = TRUE)
median_bathroom <- median(as.numeric(as.character(property_final$number_of_bathrooms)), na.rm = TRUE)

property_CT <- property_final %>%
  filter(!zoning %in% c("I2", "SPPOA"))%>%
  mutate(
        # Neighborhood Name
    nhoods_name = case_when(
      nhoods_name %in% c("CENTER_CITY", "LOGAN_SQUARE") ~ "WEST_PART",
      TRUE ~ nhoods_name
    ),

    # Land Use
    landuse = case_when(
      landuse %in% c("Park/Open Space", "Other/Unknown", "Commercial Business/Professional") ~ "Open Space/Business",
      landuse %in% c("Commercial Consumer", "Commercial Mixed Residential",
                     "Industrial", "Transportation", "Culture/Amusement") ~ "Commercial",
      landuse %in% c("Civic/Institution", "Residential High") ~ "Residential High/Civic",
      TRUE ~ landuse
    ),

    # LPSS Category
#    LPSS_PER1000 = case_when(
#      LPSS_PER1000 < 45 ~ "low",
#      LPSS_PER1000 >= 45 & LPSS_PER1000 <= 60.70941337 ~ "mid",
#      LPSS_PER1000 %in% c(60.70941337, 131.86813187) ~ "high",
#      TRUE ~ "Unknown"
#    ),
#    LPSS_PER1000 = factor(LPSS_PER1000, levels = c("low", "mid", "Unknown")),

    # HPSS Category
#    HPSS_PER1000 = case_when(
#      HPSS_PER1000 < 5.6 ~ "low",
#     HPSS_PER1000 >= 5.6 & HPSS_PER1000 <= 7.3 ~ "mid",
#      HPSS_PER1000 >= 7.3 ~ "high",
#      TRUE ~ "Unknown"
#    ),
#    HPSS_PER1000 = factor(HPSS_PER1000, levels = c("low", "mid", "Unknown")),

    # Year Built
    year_built = as.numeric(year_built),
    year_built_missing = ifelse(is.na(year_built), 1, 0),
#    year_built = case_when(
#      is.na(year_built) ~ "Unknown",
#      TRUE ~ as.character(year_built)
#    ),
#    year_built = factor(year_built),
    
    year_built_group = case_when(
     year_built < 1900 ~ "1900 Before",
     year_built >= 1900 & year_built < 1930 ~ "1900- 1930",
     year_built >= 1930 & year_built < 1970 ~ "1930 - 1970",
     year_built >= 1970 & year_built < 2010 ~ "1970 - 2010",
     year_built >= 2010 ~ "2010+",
      TRUE ~ "Unknown"
    ),
    year_built_group = factor(year_built_group, levels = c("1900 Before","1900- 1930","1930 - 1970","1970 - 2010","2010+","Unknown")),

    # Log Transforms
    log_price = log(adj_sale_price),
    log_dist_highway = log(distance_to_I676 + 1),

    # Central Air
    central_air = case_when(
      central_air %in% c("Y", "1")  ~ "Y",
      central_air %in% c("N", "0") ~ "N",
      TRUE ~ "Unknown"
    ),
    central_air = factor(central_air),
    central_air_missing = ifelse(central_air == "Unknown", 1, 0),

    # Garage Spaces
    garage_spaces_missing = ifelse(is.na(garage_spaces), 1, 0),
    garage_spaces = case_when(
#      is.na(garage_spaces) ~ "Unknown",
      garage_spaces == 0 | garage_spaces == 1 ~ "less than 2",
      TRUE ~ "2 or more"
    ),
    garage_spaces = factor(garage_spaces),

    # Bedrooms
    number_of_bedrooms = as.numeric(as.character(number_of_bedrooms)),
    number_of_bedrooms_missing = ifelse(is.na(number_of_bedrooms), 1, 0),
    number_of_bedrooms_filled = ifelse(is.na(number_of_bedrooms), median_bedroom, number_of_bedrooms),
    number_of_bedrooms = case_when(
      is.na(number_of_bedrooms) ~ "Unknown",
      number_of_bedrooms == 0 ~ "0",
      number_of_bedrooms == 1 ~ "1",
      number_of_bedrooms %in% c(2, 3) ~ "2",
      number_of_bedrooms %in% c(5, 6) ~ "3",
      number_of_bedrooms == 4 ~"4",
      number_of_bedrooms %in% c(7, 9, 12) ~ "5"
    ),
    number_of_bedrooms = factor(number_of_bedrooms),

    # Bathrooms
    number_of_bathrooms = as.numeric(as.character(number_of_bathrooms)),
    number_of_bathrooms_missing = ifelse(is.na(number_of_bathrooms), 1, 0),
    number_of_bathrooms_filled = ifelse(is.na(number_of_bathrooms), median_bathroom, number_of_bathrooms),
    number_of_bathrooms = case_when(
      number_of_bathrooms == 0 ~ "0",
      number_of_bathrooms == 1 ~ "1",
      number_of_bathrooms == 2 ~ "2",
      number_of_bathrooms %in% c(3, 4, 5) ~ "3-5",
      number_of_bathrooms == 6 ~ "6",
      number_of_bathrooms == 7 ~ "7",
      TRUE ~ "Unknown"
    ),
    number_of_bathrooms = factor(number_of_bathrooms),

    # Total Livable Area
    total_livable_area = ifelse(is.na(total_livable_area) | total_livable_area == 1, 0, total_livable_area),

    # Quality Grade
    quality_grade = case_when(
      quality_grade %in% c("A ", "A-") | is.na(quality_grade) ~ "Good",
      quality_grade %in% c("3 ", "C+", "C")  ~ "Average",
      TRUE ~ "Below Average"
    ),
    quality_grade = factor(quality_grade, levels = c("Good", "Average", "Below Average")),
    quality_grade_missing = ifelse(is.na(quality_grade), 1, 0),

    # Construction
    general_construction = as.character(general_construction),
    general_construction = case_when(
      is.na(general_construction) ~ "Unknown",
      TRUE ~ general_construction
    ),
    general_construction = factor(general_construction),
    general_construction_missing = ifelse(general_construction == "Unknown", 1, 0),

    # Separate Utilities
    separate_utilities = case_when(
      is.na(separate_utilities) ~ "Unknown",
      TRUE ~ separate_utilities
    ),
    separate_utilities = factor(separate_utilities, levels = c("A", "B", "C", "Unknown")),
    separate_utilities_missing = ifelse(separate_utilities == "Unknown", 1, 0),

    # Exterior Condition
    exterior_condition = case_when(
      exterior_condition == 2 ~ "Good",
      exterior_condition %in% c(1, 6, 7) | is.na(exterior_condition) ~ "Average",
      exterior_condition %in% c(3, 4, 5) ~ "Below Average"
 #     TRUE ~ "Unknown"
    ),
    exterior_condition = factor(exterior_condition),

    # Interior Condition
    interior_condition = case_when(
      interior_condition %in% c(1,2,7) | is.na(interior_condition) ~ "Good",
      interior_condition %in% c(3,4,5) ~ "Average",
      interior_condition %in% c(6) ~ "Below Average",
      TRUE ~ "Unknown"
    ),
    interior_condition = factor(interior_condition),

    # Category Code
    category_code_original = as.character(category_code),
    category_code_group = case_when(
      category_code == "12" ~ "6 ",
      category_code == "14" ~ "2 ",
      category_code == "10" ~ "4 ",
      category_code == "15" ~ "5 ",
      category_code == "16" ~ "4 ",
      category_code == "7 "  ~ "4 ",
      category_code == "8 "  ~ "6 ",
      category_code == "9 "  ~ "2 ",
      category_code %in% c("1 ", "2 ", "3 ", "4 ", "5 ", "6 ") ~ category_code,
      TRUE ~ "Other"
    ),
    category_code = factor(category_code_group),

    # Zoning
    zoning_group = case_when(
      zoning %in% c("CMX5") ~ "CMX5",
      zoning %in% c("IRMX") ~ "IRMX",
      zoning %in% c("RM1", "RM2", "RM4") ~ "RM",
      zoning %in% c("RMX3") ~ "RMX3",
      zoning %in% c("RSA5") ~ "RSA5",
      zoning %in% c("ICMX") ~ "ICMX",
      zoning %in% c("SPPOA") ~ "SPPOA",
      zoning %in% c("I2") ~ "I2",
      zoning %in% c("CMX2.5", "CMX2", "CMX1") ~ "CMX(1,2,2.5)",
      zoning %in% c("CMX3", "CMX4") ~ "CMX(3,4)",
      TRUE ~ "Other"
    ),
    zoning_group = factor(zoning_group),

    zoning = case_when(
      is.na(zoning) ~ "Unknown",
      TRUE ~ zoning
    ),
    zoning = factor(zoning),
    zip_code = factor(zip_code)
  )

mean_depth <- mean(property_CT$depth, na.rm = TRUE)
property_CT$depth[is.na(property_CT$depth)] <- mean_depth

st_write(property_CT, "property_CT_FE.geojson", driver = "GeoJSON", delete_dsn = TRUE)
```



### Log Transformation
```{r Log Transofrmed - Distribution of Total Livable Area - sale price, message=FALSE, warning=FALSE}
plot_df <- property_final %>%
  mutate(
    total_livable_area_log = log1p(total_livable_area) 
  )

plot1 <- ggplot(plot_df, aes(x = total_livable_area)) +
  geom_histogram(bins = 30, fill = "#1B8370", alpha = 0.8) +
  labs(title = "Original Total Livable Area",
        x = "Total Livable Area", 
        y = "Count") +
  xlim(0, quantile(plot_df$total_livable_area, 0.99, na.rm = TRUE)) + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5)
  )

plot2 <- ggplot(plot_df, aes(x = total_livable_area_log)) +
  geom_histogram(bins = 30, fill = "#1B8370", alpha = 0.8) +
  labs(title = "Log-transformed Total Livable Area",
        x = "(Log) Total Livable Area", 
        y = "Count")  +
  xlim(5, 9) +  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
combined_plot <- plot1 + plot2 +
  plot_layout(widths = c(1, 1)) +
  plot_annotation(
    title = "Distribution of Total Livable Area",
    caption = "Figure X.X",
    theme = theme(
      plot.title = element_text(size = 16,face = "bold", hjust = 0.5),
      plot.caption = element_text(hjust = 0.5)
    )
  )
combined_plot
```

### Reclassification
```{r Reclassification - interior condition - Mean Sale Price, message=FALSE, warning=FALSE}
lm1_property <- property_final %>%
  mutate(
    interior_condition_recode = case_when(
      interior_condition %in% c(1,2,7) | is.na(interior_condition) ~ "Good",
      interior_condition %in% c(3,4,5) ~ "Average",
      interior_condition %in% c(6) ~ "Below Average",
      TRUE ~ "Unknown"
    ),
    interior_condition_recode = factor(interior_condition_recode, levels = c("Good", "Average", "Below Average", "Unknown"))
  )
plot1 <- ggplot(lm1_property, aes(x = factor(interior_condition))) +
  geom_bar(fill = "#1B8370", alpha = 0.8, width = 0.6) +
  ggtitle("Original Interior Condition") +
  labs(x = "Interior Condition Code", y = "Count") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))

plot2 <- lm1_property %>%
  group_by(interior_condition) %>%
  summarise(mean_log_price = mean(market_value, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(interior_condition), y = mean_log_price)) +
  geom_col(fill = "#E46726", alpha = 0.8, width = 0.6) +
  ggtitle("Mean Sale Price by Original Condition") +
  labs(x = "Interior Condition Code", y = "Avg Sale Price ($)") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))

plot3 <- ggplot(lm1_property, aes(x = interior_condition_recode)) +
  geom_bar(fill = "#1B8370", alpha = 0.8, width = 0.6) +
  ggtitle("Reclassified Interior Condition") +
  labs(x = "Interior Condition Recode", y = "Count") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))

plot4 <- lm1_property %>%
  group_by(interior_condition_recode) %>%
  summarise(mean_log_price = mean(market_value, na.rm = TRUE)) %>%
  ggplot(aes(x = interior_condition_recode, y = mean_log_price)) +
  geom_col(fill = "#E46726", alpha = 0.8, width = 0.6) +
  ggtitle("Mean Sale Price by Reclassified Condition") +
  labs(x = "Interior Condition Recode", y = "Avg Sale Price ($)") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))

# Combine with patchwork + large centered title
combined_plot <- (plot1 + plot2 + plot_layout(tag_level = "new")) /
                 (plot3 + plot4 + plot_layout(tag_level = "new")) +
  plot_annotation(
    title = "Interior Condition and Sale Price",
    subtitle = "Before (Top) and After (Bottom) Reclassification",
    caption = "Figure X.X",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      plot.caption = element_text(hjust = 0.5),
      plot.title.position = "plot"
    )
  )
combined_plot
```


## Comparing North and South - Continuous Variables
```{r message=FALSE, warning=FALSE}
property_fe <- property_CT %>%
  st_drop_geometry() %>%
  mutate(
    log_sale_price = log1p(adj_sale_price),
    log_distance_to_I676 = log1p(distance_to_I676),
    log_distance_to_city_hall = log1p(distance_to_city_hall),
    log_total_livable_area = log1p(total_livable_area),
    I676_NS = as.character(I676_NS) 
  )

property_fe_selected <- property_fe %>%
  select(
    I676_NS,
    log_sale_price,
    log_distance_to_I676,
    log_distance_to_city_hall,
    log_total_livable_area
  ) %>%
  drop_na()

correlation_long <- property_fe_selected %>%
  pivot_longer(-c(log_sale_price, I676_NS), names_to = "Variable", values_to = "Value") %>%
  mutate(
    use_in_plot = case_when(
      Variable == "log_total_area" & Value <= log1p(10) ~ FALSE,
      Variable == "log_total_livable_area" & Value <= log1p(10) ~ FALSE,
      TRUE ~ TRUE
    )
  )

cor_selected <- correlation_long %>%
  filter(use_in_plot) %>%
  group_by(Variable, I676_NS) %>%
  summarise(r = cor(Value, log_sale_price, use = "complete.obs"), .groups = "drop") %>%
  mutate(
    x_pos = -Inf,
    y_pos = ifelse(I676_NS == "north", Inf, -Inf),
    vjust_pos = ifelse(I676_NS == "north", 1.1, -1.1)
  )

ggplot(correlation_long %>% filter(use_in_plot),
       aes(x = Value, y = log_sale_price)) +
  geom_point(aes(color = I676_NS), alpha = 0.4, size = 0.3) +
  geom_smooth(aes(color = I676_NS), method = "lm", se = FALSE, linewidth = 1) +
  facet_wrap(~ Variable, scales = "free", ncol = 3) +
  scale_color_manual(values = c("north" = "#1a9988", "south" = "#eb5600")) +
  labs(
    title = "Log-transformed Sale Price vs Continuous Features by I-676 Side",
    caption = "Figure X.X",
    x = "Values of Continuous Features",
    y = "Log(Sale Price)",
    color = "I-676 Side"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 15, face = "bold",hjust = 0.5),
    strip.text = element_text(size = 10, face = "bold"),
    plot.caption = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

```



## Comparing North and South - Categorical Variables
```{r}
# Prepare aggregated data by zoning group and I676 side
zoning_bar_data <- property_CT %>%
  st_drop_geometry() %>%
  mutate(I676_NS = factor(I676_NS, levels = c("north", "south")))%>%
  filter(!is.na(zoning_group), !is.na(adj_sale_price), !is.na(I676_NS)) %>%
  group_by(zoning_group, I676_NS) %>%
  summarise(mean_sale_price = mean(adj_sale_price, na.rm = TRUE), .groups = "drop")

# Prepare aggregated data by quality grade and I676 side
quality_bar_data <- property_CT %>%
  st_drop_geometry() %>%
  mutate(I676_NS = factor(I676_NS, levels = c("north", "south")))%>%
  filter(!is.na(quality_grade), !is.na(adj_sale_price), !is.na(I676_NS)) %>%
  group_by(quality_grade, I676_NS) %>%
  summarise(mean_sale_price = mean(adj_sale_price, na.rm = TRUE), .groups = "drop")
base_font_size <- 10

plot_zoning <- ggplot(zoning_bar_data, aes(x = zoning_group, y = mean_sale_price, fill = I676_NS)) +
  geom_col(position = "dodge", alpha = 0.9) +
  scale_fill_manual(values = c("north" = "#1a9988", "south" = "#eb5600")) +
  labs(
    title = "Average Sale Price by Zoning Group",
    x = "Zoning Group", y = "Avg Sale Price ($)"
  ) +
  theme_minimal(base_size = base_font_size) +
  theme(
    plot.title = element_text(face = "bold", size = base_font_size + 2, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = base_font_size - 1),
    legend.position = "none"
  )

plot_quality <- ggplot(quality_bar_data, aes(x = factor(quality_grade), y = mean_sale_price, fill = I676_NS)) +
  geom_col(position = "dodge", alpha = 0.9) +
  scale_fill_manual(values = c("north" = "#1a9988", "south" = "#eb5600")) +
  labs(
    title = "Average Sale Price by Quality Grade",
    x = "Quality Grade", y = "Avg Sale Price ($)",
    fill = "I-676 Side"
  ) +
  theme_minimal(base_size = base_font_size) +
  theme(
    plot.title = element_text(face = "bold", size = base_font_size + 2, hjust = 0.5),
    axis.text.x = element_text(size = base_font_size - 1)
  )

final_plot <- plot_zoning + plot_quality +
  plot_layout(ncol = 2,  guides = "collect") 

final_plot
```


```{r Zoning, message=FALSE, warning=FALSE, quiet=TRUE, results='hide', eval = FALSE}
zoning_group_colors <- c(
  "RM" = "#F9844A",                 
  "RSA5" = "#E89972",             
  "RMX3" = "#BFD3C1",             
  "CMX5" = "#3B7D64",             
  "ICMX" = "#004B40",              
  "IRMX" = "#FFDAB9",             
  "RMX1" = "#F9C74F",             
  "SPPOA" = "#F8961E", 
  "CMX(1,2,2.5)" = "#43AA8B",
  "CMX(3,4)" = "#90BE6D",
  "Other" = "grey80"
)

ggplot() +
  geom_sf(data = block2023_studyarea, fill = "grey95", color = "grey80", size = 0.3) +
  geom_sf(data = I676_2, color = "red", size = 1, alpha = 0.9) +
  geom_sf(data = property_CT %>% filter(!is.na(zoning_group)),
          aes(color = zoning_group), size = 1, alpha = 0.7) +
  scale_color_manual(values = zoning_group_colors, na.translate = FALSE) +
  labs(
    title = "Zoning Groups of Properties within Study Area",
    subtitle = "Study Area Around I676",
    caption = "Figure X.X",
    color = "Zoning Group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(hjust = 0.5),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  )
```

```{r LPSS, message=FALSE, warning=FALSE, quiet=TRUE, results='hide', eval = FALSE}
property_CT_LPSS <- property_CT %>%
  mutate(
    LPSS_CAT = case_when(
      LPSS_PER1000 < 45 ~ "low",
      LPSS_PER1000 >= 45 & LPSS_PER1000 < 60 ~ "mid",
      LPSS_PER1000 >= 60 ~ "high",
      TRUE ~ "Unknown"
    ),
    LPSS_CAT = factor(LPSS_CAT, levels = c("low", "mid", "high", "Unknown"))
  )
ggplot() +
  geom_sf(data = block2023_studyarea, fill = "grey95", color = "grey80", size = 0.3) +
  geom_sf(data = I676_2, color = "red", size = 1, alpha = 0.9) +
  geom_sf(
    data = property_CT_LPSS %>% filter(!is.na(LPSS_PER1000)),
    aes(color = LPSS_CAT),  
    size = 1,
    alpha = 0.8
  ) +
  scale_color_manual(
    values = c(
      "low" = "#FFDDAB",
      "mid" = "#E89972",
      "high" = "#004B40",
      "Unknown" = "grey70"
    ),
    name = "LPSS Level"
  ) +
  labs(
    title = "Low Produce Supply Score (LPSS) within Study Area",
    subtitle = "Darker = Higher LPSS, i.e., More Limited Access",
    caption = "Figure X.X"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(hjust = 0.5),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  )
```

# Model
## OLS Model
```{r message=FALSE, warning=FALSE}
lm_CT <- lm(
  log(adj_sale_price) ~ 
    log_dist_highway + 
    log(distance_to_city_hall+1) + 
#    distance_to_nearest_transit + 
#    distance_to_nearest_school + 
#    distance_to_nearest_hospital +
#    log(distance_to_nearest_bikelane) +
    log(nearest_restaurant_dist_m + 1) + 
    log(distance_to_nearest_park + 1) + 
    zoning_group + 
    zip_code + 
    landuse+
    nhoods_name+
    exterior_condition + 
    interior_condition + 
    log(total_livable_area + 1) + 
    number_of_bathrooms + 
    number_of_bedrooms + 
    quality_grade + 
    log(distance_to_nearest_water + 1) + 
    central_air + 
    year_built_group + 
 #   year_built_missing + 
    LPSS_PER1000+
#    HPSS_PER1000+
    general_construction + 
    garage_spaces + 
    separate_utilities + 
#    central_air_missing +
#    category_code + 
    crime_nn5,
#    general_construction_missing + 
#    number_of_bathrooms_missing + 
#    number_of_bedrooms_missing + 
#    separate_utilities_missing,
  data = property_CT
)

model_summary <- tidy(lm_CT) %>%
  mutate(
    p.value = signif(p.value, 3),
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 2),
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    )
  ) %>%
  select(term, estimate, std.error, statistic, p.value, sig)
predicted <- predict(lm_CT)
actual <- log(property_CT$adj_sale_price)
rmse <- sqrt(mean((predicted - actual)^2, na.rm = TRUE))

glance_stats <- glance(lm_CT)
model_stats <- tibble(
  term = c("R-squared", "Adj. R-squared", "Num. Obs.", "RMSE"),
  estimate = c(
    round(glance_stats$r.squared, 3),
    round(glance_stats$adj.r.squared, 3),
    glance_stats$nobs,
    round(rmse, 3)
  ),
  std.error = NA, statistic = NA, p.value = NA, sig = ""
)
model_full <- bind_rows(model_summary, model_stats)

model_full %>%
  kable("html", caption = "OLS Regression Results", digits = 3, align = "c") %>%
  kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped", "hover", "condensed"),
    position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = "#1a9988", color = "white") %>%
  scroll_box(height = "350px") 
```


```{r}
feature_lookup <- tibble(
  Variable = c(
    "log_dist_highway",
    "log(distance_to_city_hall + 1)",
    "log(nearest_restaurant_dist_m + 1)",
    "log(distance_to_nearest_park + 1)",
    "zoning_group",
    "zip_code",
    "exterior_condition",
    "interior_condition",
    "log1p(total_livable_area)",
    "number_of_bathrooms",
    "number_of_bedrooms",
    "quality_grade",
    "log(distance_to_nearest_water + 1)",
    "central_air",
    "year_built_group",
    "LPSS_PER1000",
    "general_construction",
    "garage_spaces",
    "separate_utilities",
    "category_code",
    "crime_nn5",
    "landuse",
    "nhoods_name"
  ),
  Feature = c(
    "Log Distance to Highway",
    "Log Distance to City Hall",
    "Log Distance to Nearest Restaurant",
    "Log Distance to Nearest Park",
    "Zoning Group",
    "Zip Code",
    "Exterior Condition",
    "Interior Condition",
    "Log Total Livable Area",
    "Number of Bathrooms",
    "Number of Bedrooms",
    "Quality Grade",
    "Log Distance to Nearest Water",
    "Central Air",
    "Year Built",
    "Low-Produce Supply Stores per 1000",
    "General Construction",
    "Garage Spaces",
    "Separate Utilities",
    "Category Code",
    "Crime Rate (NN5)",
    "Land Use",
    "Neighborhood"
  )
)

vif_df <- car::vif(lm_CT) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  as_tibble() %>%
  rename(
    GVIF = `GVIF`,
    Df = `Df`,
    VIF = `GVIF^(1/(2*Df))`
  ) %>%
  mutate(
    VIF = round(VIF, 2),
    Level = case_when(
      VIF >= 10 ~ "Very High",
      VIF >= 5 ~ "High",
      VIF < 5 ~ "OK",
      TRUE ~ "Unknown"
    )
  )

vif_table <- vif_df %>%
  left_join(feature_lookup, by = "Variable") %>%
  select(Variable, Feature, VIF, Level)
vif_table %>%
  kable("html", caption = "VIF (GVIF-adjusted) with Feature Descriptions", align = "c") %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed"),
    position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = "#1a9988", color = "white") %>%
  row_spec(which(vif_table$VIF >= 5), background = "#eb5600") %>%
  scroll_box(height = "300px")
```

## OLS predictive model
```{r message=FALSE, warning=FALSE, quiet=TRUE, results='hide'}
set.seed(123)
split <- initial_split(property_CT, prop = 0.8)  # 80% train, 20% test

train_data <- training(split)
test_data <- testing(split)
```

```{r message=FALSE, warning=FALSE, quiet=TRUE, results='hide'}
lm_train <- lm(
  log(adj_sale_price) ~ 
    log_dist_highway + 
    log(distance_to_city_hall+1) + 
#    distance_to_nearest_transit + 
#    distance_to_nearest_school + 
#    distance_to_nearest_hospital +
#    log(distance_to_nearest_bikelane) +
    log(nearest_restaurant_dist_m + 1) + 
    log(distance_to_nearest_park + 1) + 
    zoning_group + 
    zip_code + 
    exterior_condition + 
    interior_condition + 
    log1p(total_livable_area) + 
    distance_to_nearest_hospital + 
    number_of_bathrooms + 
    number_of_bedrooms + 
    quality_grade + 
    log(distance_to_nearest_water + 1) + 
    central_air + 
    year_built_group + 
    nhoods_name+
    landuse+
#    year_built_missing + 
    LPSS_PER1000+
#    HPSS_PER1000+
    general_construction + 
    garage_spaces + 
    separate_utilities + 
    category_code + 
    crime_nn5,
  data = train_data
)
```

```{r message=FALSE, warning=FALSE, quiet=TRUE}

test_data$log_pred <- predict(lm_train, newdata = test_data)
test_data$price_pred <- exp(test_data$log_pred) - 1

test_data <- test_data %>%
  mutate(
    abs_error = abs(price_pred - adj_sale_price),
    pct_error = abs_error / adj_sale_price
  )
#mean(test_data$pct_error, na.rm = TRUE) 
#median(test_data$pct_error, na.rm = TRUE) 

ggplot(test_data, aes(x = price_pred, y = adj_sale_price)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Test Set: Predicted vs Actual Sale Price",
    x = "Predicted Price",
    y = "Actual Price"
  ) +
  theme_minimal()
rmse_value <- sqrt(mean((test_data$price_pred - test_data$adj_sale_price)^2, na.rm = TRUE))
mae_value <- mean(abs(test_data$price_pred - test_data$adj_sale_price), na.rm = TRUE)

```

```{r message=FALSE, warning=FALSE, quiet=TRUE, results='hide', eval = FALSE}
property_predict <- property_CT %>%
  mutate(
    log_price = log(adj_sale_price),
    log_pred = predict(lm_train, newdata = .)
  )
st_write(property_predict, "property_pred_OLS.geojson", driver = "GeoJSON", delete_dsn = TRUE)
comparison_df <- property_predict %>%
  select(location, log_price, log_pred)
```



